---
title: "Data Analysis Exercise"
author: "Leah Lariscy"
editor: visual
---

## Load libraries here

```{r}
library(tidyverse)
library(lubridate)
```

## Data description

This dataset was accessed from the **Bacteria, Enterics, Amoeba, and Mycotics (BEAM) Dashboard**, which houses data collected by the **System for Enteric Disease Response, Investigation, and Coordination (SEDRIC)**. The data points represent pathogens isolated from infected individuals, namely ***Salmonella***, ***E. Coli***, ***Shigella***, and ***Campylobacter*** bacteria.

You can access the BEAM Dashboard [here](https://www.cdc.gov/ncezid/dfwed/beam-dashboard.html).

```{r}
beam_raw <- read_csv("data_analysis_exercise/data/raw_data/BEAM_Dashboard_Report_Data.csv")
```

## Data cleaning

Begin by combining the `Year` and `Month` variables into one `Date` variable, then select for variables of interest and rename them. Lastly, change all NAs to 0s where necessary.

```{r}
beam_clean <- beam_raw %>% mutate(date = make_date(year=Year, month=Month, day=1)) %>% 

  select(c(date, state=State, source=Source, pathogen=Pathogen, species=`Serotype/Species`,
       n_isolates=Number_of_isolates, n_outbreak_associated=Outbreak_associated_isolates)) %>% 

  mutate(n_isolates = ifelse(is.na(n_isolates), 0, n_isolates), 
       n_outbreak_associated = ifelse(is.na(n_outbreak_associated), 0, n_outbreak_associated))

print(beam_clean)
```

## Save cleaned data to RDS file
```{r}
saveRDS(beam_clean, "data_analysis_exercise/data/raw_data/BEAM_Report_clean.RDS")
```

## Read in RDS file 
```{r}
#read_rds("data_analysis_exercise/data/raw_data/BEAM_Report_clean.RDS")
```

## Exploratory analysis
I would be interested in grouping by pathogen and exploring trends across states. I'd also like to see which pathogens are more likely to be associated with outbreaks. 