---
title: "Data Analysis Exercise"
author: "Leah Lariscy"
editor: visual
---

## Load libraries here

```{r}
library(tidyverse)
library(lubridate)
```

## Data description

This dataset was accessed from the **Bacteria, Enterics, Amoeba, and Mycotics (BEAM) Dashboard**, which houses data collected by the **System for Enteric Disease Response, Investigation, and Coordination (SEDRIC)**. The data points represent pathogens isolated from infected individuals, namely ***Salmonella***, ***E. Coli***, ***Shigella***, and ***Campylobacter*** bacteria.

You can access the BEAM Dashboard [here](https://www.cdc.gov/ncezid/dfwed/beam-dashboard.html).

```{r}
beam_raw <- read_csv("data_analysis_exercise/data/raw_data/BEAM_Dashboard_Report_Data.csv")
```

## Data cleaning

Begin by combining the `Year` and `Month` variables into one `Date` variable, then select for variables of interest and rename them. Lastly, change all NAs to 0s where necessary.

```{r}
beam_clean <- beam_raw %>% mutate(date = make_date(year=Year, month=Month, day=1)) %>% 

  select(c(date, state=State, source=Source, pathogen=Pathogen, species=`Serotype/Species`,
       n_isolates=Number_of_isolates, n_outbreak_associated=Outbreak_associated_isolates)) %>% 

  mutate(n_isolates = ifelse(is.na(n_isolates), 0, n_isolates), 
       n_outbreak_associated = ifelse(is.na(n_outbreak_associated), 0, n_outbreak_associated))

print(beam_clean)
```

## Save cleaned data to RDS file

```{r}
saveRDS(beam_clean, "data_analysis_exercise/data/raw_data/BEAM_Report_clean.RDS")
```

## Read in RDS file

```{r}
#read_rds("data_analysis_exercise/data/raw_data/BEAM_Report_clean.RDS")
```

## Exploratory analysis

I would be interested in grouping by pathogen and exploring trends across states. I'd also like to see which pathogens are more likely to be associated with outbreaks.

## This section was added by Kimberly Perez

#Reading in RDS

Here I will use the readRDS() function to load Leah's cleaned BEAM data.

```{r}
#Utilize readRDS for this task
newbc<-readRDS("data_analysis_exercise/data/raw_data/BEAM_Report_clean.RDS")
```

## A bit more data wrangling

I will now select four states to analyze

```{r}
#Creating a new dataframe with four selected states
bc<- newbc %>%
  filter(state == 'GA'  | state == 'CA'| state == 'ND' | state == 'OR' )
```

# Selecting and condensing data for visualization!

Next, I will use the filter function to select four states, remove "n_isolate" and "species" columns, and find count information on each pathogen. I will then produce a visualization that will present case count information for each pathogen by state throughout 2017-2022. Hopefully, this will provide us with some basic trend information.

```{r}
#Creating a new dataframe, grouping variables, and removing n_isolates, source, n_outbreak_associated, and species columns 
bc_path<- bc %>% 
  group_by(date, pathogen, state) %>% 
  mutate(count=n()) %>% 
  ungroup() %>%  
  select(-c(n_isolates, species, source, n_outbreak_associated))

#Removing duplicates (by mean) and creating a new dataframe presenting case counts for each pathogen in each state by day (2017-2022)
final_bcc<- bc_path %>% 
  group_by(date,state,pathogen) %>% 
  summarize_if(is.numeric, mean) %>% 
  ungroup()

#Graphing
final_bcc %>% 
  ggplot() +geom_line(
  aes(x = date,
      y = count,
      color = pathogen,
      linetype = state)) +
  theme_bw() +
  labs(x = "Year",
       y = "Case Count",
       color= "Pathogen",
       linetype="State",
       title = "Pathogen trends from 2017-2022") +
  theme(plot.title = element_text(hjust = 0.6))

```

From the graph rendered above, it seems as though ***Salmonella*** comprises the majority of reported cases for most states. 

## Visualizing Outbreak by State and Pathogen (2017-2022) in two ways...
```{r}
theme_set(theme_bw())  

#Visualizing the data...
bc1 <- ggplot(bc, aes(state, n_outbreak_associated)) + 
  labs(title="Pathogen Associated with Outbreak by State (2017-2022) by State",
      y= "Number of Outbreaks",
      x= "State")

bc1 + geom_jitter(aes(col=pathogen, size=n_outbreak_associated)) + 
  geom_smooth(aes(col=pathogen), method="lm", se=F)


#Another way to visualize the data...
bc %>% 
  ggplot() +geom_line(
  aes(x = date,
      y = n_outbreak_associated,
      color = pathogen,
      linetype = state)) +
  theme_bw() +
  labs(x = "Year",
       y = "Number of Outbreaks (Associated with Pathogen)",
       color= "Pathogen",
       linetype="State",
       title = "Pathogen Associated with an Outbreak by State (2017-2022)") +
  theme(plot.title = element_text(hjust = 0.6))
```
***Salmonella*** and ***Ecoli*** look to be the pathogens that are associated with the largest number of outbreaks among these four states.

### Pathogen Source by State

Have you ever wondered what laboratories isolate pathogens from? Let's explore the possibilities below!
```{r}
#Plotting 
ggplot(bc, aes(state, pathogen, colour = source)) + geom_count(show.legend=T) +
  labs(y="Pathogen", 
       x="State", 
       title="Pathogen Source by State (2017-2022)")
```
From the visualization above, it looks like most labs are able to isolate pathogens from **stool** samples, however, there may be variation by state!

## Potentially helpful resource
[This website](http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html#Counts%20Chart) seems to be a helpful resource for data visualization. I found some interesting visualizations and hope to practice utilizing several.
