---
title: "Flu Data Wrangling"
author: "Leah Lariscy"
editor: visual
format:
  html:
    toc: true
    toc-depth: 3
---

# Wrangling time..

## Load libraries

```{r}
library(tidyverse)
library(here)
```

## Load data

```{r}
symptoms <- readRDS(here("fluanalysis/data/raw_data/SympAct_Any_Pos.Rda"))
```

## Look at data

```{r}
glimpse(symptoms)
skimr::skim(symptoms)
```

After viewing the data, I see there are 63 variables and 735 observations. Most are coded as factors and integers, and there is 1 character variable. Some variables have quite a good amount of NAs.

## Remove unnecessary variables

```{r}
symptoms <- symptoms %>% select(9:40) %>% select(!contains("YN", ignore.case = FALSE))
```

Since all the variables of interest were all consecutive, I was able to easily select them based on their range of column number.

I then selected for all variables that did not include "YN", removing duplicate variables.

### Remove "near-zero" variables with less than 50 entries in one binary category

```{r}
summary(symptoms) #Vison and Hearing have the less than 50 yes

symptoms <- symptoms %>% select(!c(Vision, Hearing))
```

## Remove NAs

```{r}
symptoms <- symptoms %>% na.omit()
```

Only 5 observations were removed, so it looks like we took care of most of the NAs when we selected for the relevant variables.

## Cleaning is complete, save as RDS file

```{r}
saveRDS(symptoms, here("fluanalysis/data/processed_data/symptoms_clean.RDS"))
```
